---
title: "Reproducible Report Template - Basic"
author: "Student/s SID"
subtitle: "Project 1"
date: "University of Sydney | Data1001 | August 2018"
# You can change the title, subtitle, author etc.
# if you do not want a subtitle, author, or date just delete or comment # the the line!
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    # Other themes can be found at: https://bootswatch.com/
    css: 
      - https://use.fontawesome.com/releases/v5.0.6/css/all.css
    #This will allow you to insert icons in your document, thanks to [Font Awesome](http://fontawesome.io).
    #You will find an examples below.
    # If you are knitting your document but not connected to the internet, place a "#" in front of the css line above with the font-awesome.min.css line to comment it out. This will let you knit the document in draft form until you connect again and knit the final. Note that when you do this some elements will be missing from the knitted slides. 
    toc: true
    #This adds a Table of Contents (toc).
    toc_depth: 3
    #This controls the number of subheadings included in the toc.
    toc_float: true
    #This allows toc to update relative to your location as you scroll down the html page.
    code_folding: hide
    #code_folding can be changed to "hide"" if you want the all your code to be collapsed when you 
---

<br>

# Executive Summary
- The aim of this report is to find out what kind of correlations are between weather and alcohol sales in Sydney.
- The main discoveries are ...

<br>

<br>
# Full Report

## Initial Data Analysis (IDA)
### Source of Data
This report uses data from an off-premise liquor store called The Drink Hive in a small up-market shopping centre called Saporium in Rosebery, NSW Australia. 

#### Rosebery Demographic
Rosebery has a population of approximately 10,000 with a median age of 33 and 2,500 families. The median weekly income is $1,900 and weekly rent is $580. 34% of residents have a university or tertiary education which is twice that of NSW average and 67% work full-time. 58% of the population live in flats or apartments.
The suburb has a diverse mix of cultures with Chinese and Greek ancestory heavily over-represented compared to the NSW Average. 63% of the population's parents were both born overseas, just under double the NSW average. Eastern orthodox over-represented and Anglican under-represented.(ABS, 2016)

### Domain Knowledge
This report looks at the effect of weather on alcohol sales in the hope of improving the likelihood of small business success. A business's core role is to provide goods and services. Their success is dependent on their ability to sell products for revenues. It is therefore imperative that they
  1. Stock the correct product
  2. Market the correct product
  3. Price their products effectively
In the past, these decisions have been the intuition of a good business operator, however with the accessibility of data and analytics tools, we can now take a more scientific approach which can be replicated by any business.

There is a wealth of evidence which shows that weather influences consumer behaviour and understanding this leads to better marketing decisions (Murray et al, 2010). The key areas of consumer behaviour this report will address are what conditions weather effects consumer spending and what conditions effect consumer behaviour. 

Weather is defined by Merriam  as <i>the state of the atmosphere with respect to heat or cold, wetness or dryness, calm or storm, clearness or cloudiness.<\i> Researchers indicate that weather alters the shoppers mood. 

Research indicates that weather can effect consumer spending in three ways
  1. Bad weather keeps people at home reducing foot traffic and sales
  2. Weather can influence store traffic and sales volume, and
  3. Weather can influence sales by affecting the customers internal state.

We will use this report to show small businesses how they can use their data to increase profitability by making informed decisions about stocking, marketing and pricing their products, helping them find the signal in the noise.


```{r}
#This prevents strings from being factorised which keeps Date as Date
options(stringsAsFactors = FALSE)
# LOAD DATA
library(ggplot2)
library(tidyr)
data = read.csv("data/ProcessedData2.csv")

head(data$Date)
#Convert date to date format
# convert date column to date type and format to Australian date standards
data$Date <- format(as.Date(data$Date), "%d-%m-%Y")

# view R class of data
class(data$DATE)
## [1] "Date"

# view results
head(data$DATE)
## [1] "2013-08-21" "2013-08-26" "2013-08-27" "2013-09-01" "2013-09-09"
## [6] "2013-09-10"
# Quick look at top 6 rows of data
head(data)

## Size of data
dim(data)

## R's classification of data
class(data)
## R's classification of variables
str(data)
#sapply(mtcars, class)
```
### Classification of Variables
The dataset contains `r nrow(data) - 1` sales where each row represents 1 sale.
We are representing the population using a small sample described above.
We have removed columns that are not relevant to our clients use. We are using a multivariant dataset that consists of `r ncol(data)`.
Of `r names(data)`
For our research questions we will be looking at:
  "Date", as an ordinal categorical measurement representing days in a week, months in a year and     seasons in a year. The min is 
  "Total", 
 

### Stakeholders
#### Customer - Indirect External Stakeholder
The owner of The Drink Hive, Mal Higgs is a customer. He will receive this report and be able to act on its recommendations. Mal has over 40 years experience in retail liquor and therefore has a strong intuition and has made his own hypotheses/predictions as to the outcome. He came to us as he does not have a mathematical nor statistical background and his Point Of Sale (POS) software does not help with his marketing campaign. Mal is a useful resource for questions about retail and the customer however we must take care not to be bias and assess the data objectively. We must also present the report in a way that a non-technical person can understand.

#### Project Team - Direct Internal Stakeholder
The project tea is made up of Anthony, Science Major, Sindi, Ivan and Sam who are Computer Science majors. We have decided to break up the team into their strenghts and backgrounds whilst also cross-checking each other to ensure that every member of the team learns and we can challenge each others thinking. Each member of the project team will be working on the research questions presented as well as focusing on the below areas of the report.

Sam has a background in business and data analytics. He will focus on the Initial Data Analysis. He must be aware of the limitations of his understanding of data science to ensure the relevant variables are used. 
Anthony has a strong critical thinking mind and a background in scientific literature. He will focus on the evidence based conclusions, compilation and presentation of the report. He must be must be aware of his non-technical audience to present the findings in a way that can be understood by a small business owner.
Sindri and Ivar have backgrounds in Computer Science. They will be combining and wrangling the datasets into a useable table and lead the data representations found in this report. They will be ensuring the numerical and graphical summaries are insightful and relevant. They must ensure the correct data points and graphical representations are used to portray the questions being researched.

We have a fantastic breadth of knowledge that will make this report a valuable and useful tool to Mal.

#### Sponsor - Primary & Secondary Stakeholder
Our tutor and the University of Sydney are our primary and Secondary stakeholder. They assist in our vision and are responsible for the overall success of the project. They will review our submission and have created specific outcomes that we will be measured against. We must be careful to meet the requirements for this stakeholder or they will fail us and we will not be able to continue to report and continue working. 

## Summary

- The data came from two different sources. The weather data came in two different data sets from the [Bureau of Meteorology] (http://www.bom.gov.au/climate/data/) website and the sale data of alcoholic beverages came from a local store in Sydney. We are using data from both resources and combining the into one big dataset. a
- The data is/is not valid because 
  The data is valid because we are getting 
- Possible issues include ...
- Each row represents ...
- Each column represents ...

<br>

##IDA
Complexity of data:
 We are looking at a dataset with len(names())

## Research Question 1

How does the maximum temperature affect the consumer decision when purcahsing alcohol?

Insert text and analysis.

Lets start at looking at how much a person spends in the store on average. The number we see inside each bar represents the number of transactions for that temperature category and the line in each bar represent the median of money spent. As we can see from the chart, people seem to spend more money in milder temperature (15 - 40 degrees) but in more extreme temperatures (10 - 15 and 40 - 45 degrees) people tend to buy less alcohol. We can also observe that people tend to spend more money when the temperature is between 35 and 40 degrees. 



```{r}
#The temperature is a quantitative variable. We start by changing it to a qualitative one using ranges that cover 5 degrees Celcius
temp = data$Maximum.temperature..Degree.C.
data$tempGroups = cut(temp, c(10,15,20,25,30,35,40,45))

#Transaction sizes for each temperature range
meanPerPerson = aggregate(data$Total ~ data$tempGroups, data, mean)
medPerPerson = aggregate(data$Total ~ data$tempGroups, data, median)
transactions = merge(x = meanPerPerson, y = medPerPerson, by='data$tempGroups')
names(transactions) = c('Temperature', 'Mean_total', 'Median_total')
print(transactions)

Fre <- as.data.frame(table(data$tempGroups))
colnames(Fre)[1] <- "tempGroups" 
Fre$lab <- as.character(Fre$Freq)


#Barplot for average money spent with median lines
ggplot(transactions, aes(Temperature, Mean_total), label = Fre$Freq) + geom_bar(stat="identity", position = "dodge", fill = "#FF6666") + ggtitle("Average how much each person spends (in dollars) for different temperature") + ylab("Mean") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + geom_errorbar(data=transactions, aes(Temperature, ymax = Median_total, ymin = Median_total), size=1, linetype = "solid", inherit.aes = F, width = 0.9) + geom_text(aes(label = Fre$Freq), position = position_dodge(width = 0.9), vjust = 1.5) + scale_x_discrete(labels = c('10 - 15','15 - 20','20 - 25', '25 - 30', '30 - 35', '35 - 40', '40 - 45'))

 #Barplot showing only median
 #ggplot(transactions, aes(Temperature, Median_total)) + geom_bar(stat="identity", position = "dodge", fill = "#56B4E9") + ggtitle("Median how much each person spends (in dollars) for different temperature") + ylab("Median") + theme_bw() + theme(plot.title = element_text(hjust = 0.5))
 
 #ggplot(data, aes(Maximum.temperature..Degree.C., median(Total)), group = 1) + geom_boxplot() + coord_flip()

 #Number of transactions per temperature range
 #barplot(table(data$tempGroups))

```
Lets look at what an average day looks like in total sales for each temperature range. This seems to back up what we where saying before. That in extreme temperature (10 - 15 and 40 - 45 degrees) people seem to buy less alcohol while in milder temperatures people seem to buy alot. We also see the spike again when the temperature is between 35 - 40 degrees. 

```{r}
 
 #Total money spent for each temperature range
 totalPerDay = aggregate(data$Total ~ data$tempGroups, data, sum)
 nrOfDaysPerTemp = aggregate(data$Date ~ data$tempGroups, data, function(x) length(unique(x)))
 totals = merge(x = totalPerDay, y = nrOfDaysPerTemp, by='data$tempGroups')
 names(totals) = c('Temperature', 'Total', 'NrOfDays')
 totals['meanPerDay'] = round(totals$Total / totals$NrOfDays, 1)
 
 ggplot(totals, aes(Temperature, meanPerDay)) + geom_bar(stat="identity", position = "dodge") + geom_bar(stat="identity", position = "dodge", fill = "#56B4E9") + ggtitle("Average how much is spent in one day (in dollars) for different temperature") + ylab("Mean") + theme_bw() + theme(plot.title = element_text(hjust = 0.5)) + scale_x_discrete(labels = c('10 - 15','15 - 20','20 - 25', '25 - 30', '30 - 35', '35 - 40', '40 - 45'))
 
```

Summary: 
Looking at the values of the median and mean purchase transactions we see that there is not much change in consumer behaviour over the temperature ranges 15-35 degrees. However the more extreme temperature values have more of an effect. During the coldest times (10-15 degrees) there is a definate drop in amount of money spent per purchase. During the very hottest periods 40-45 degrees there is also a massive drop in amount spent on each transaction. However it is worth noting that there were very few transactions during that time. Another very interesting spike in sales occured at the 35-40 temperature range. This could be because people drink more alcohol, however these would be regular temperatures during Christamas time when people are on vacation and drink more alcohol in general.

## Research Question 2
How does rainfall affect the consumer decision when purchasing alcohol?


Insert text and analysis.

There are a few days for which we do not have rain data so we start by removing those rows. We then take a better look at the rain data Let's start by taking a better look at the rainfall data
```{r}
rainData = data %>% drop_na(Rainfall.amount..millimetres.)
print(paste('Number of rows with missing values for rainfall: ', nrow(data)-nrow(rainData)))
rain = rainData$Rainfall.amount..millimetres.
print(paste('Maximum rainfall', max(rain)))
print(paste('Minimum rainfall', min(rain)))
print(paste('Mean rainfall', mean(rain)))
print(paste('Median rainfall', median(rain)))
```

```{r}
#We start by changing the rainfall from a quantitative variable to a qualitative one
rainData$rainGroups = cut(rain, c(0,5,15,40,70))

#Transaction sizes for each temperature range
meanPerPerson = aggregate(rainData$Total ~ rainData$rainGroups, rainData, mean)
medPerPerson = aggregate(rainData$Total ~ rainData$rainGroups, data, median)
transactions = merge(x = meanPerPerson, y = medPerPerson, by='rainData$rainGroups')
names(transactions) = c('Rainfall', 'Mean_total', 'Median_total')
print(transactions)

 ggplot(transactions, aes(Rainfall, Mean_total)) + geom_bar(stat="identity", position = "dodge") 
 ggplot(transactions, aes(Rainfall, Median_total)) + geom_bar(stat="identity", position = "dodge")

 #Number of transactions for each rainfall range
 barplot(table(rainData$rainGroups))
 
 
 #Total money spent for each temperature range
 totalPerDay = aggregate(rainData$Total ~ rainData$rainGroups, rainData, sum)
 nrOfDaysPerRain = aggregate(rainData$Date ~ rainData$rainGroups, rainData, function(x) length(unique(x)))
 totals = merge(x = totalPerDay, y = nrOfDaysPerRain, by='rainData$rainGroups')
 names(totals) = c('Rainfall', 'Total', 'NrOfDays')
 totals['meanPerDay'] = round(totals$Total / totals$NrOfDays, 1)
 
 ggplot(totals, aes(Rainfall, meanPerDay)) + geom_bar(stat="identity", position = "dodge")

```



Insert text and analysis. 

Summary: 

## Research Question 3

How does the time of year affect the consumer decision when purchasing alcohol?

Other possible research questions: 
What time of day do people buy their alcohol?

Insert text and analysis. 

Summary: 

<br>

## Research Question 4

TODO:

# Conclusions

Insert text.

<br>

# References

Australian Bureau of Statistics, 2016, 2016 Census QuickStats, Rosebery. viewed 21/08/2018 http://quickstats.censusdata.abs.gov.au/census_services/getproduct/census/2016/quickstat/SSC13405?opendocument

Kyle B. Murray, Fabrizio Di Muro, Adam Finn, Peter Popkowski Leszczyc, 2010. The effect of weather on consumer spending. Journal of Retailing and Consumer Services (17) 512-520

Weather. (n.d) In Merriam-Websters Collegiate Dictionary. Retrieved from https://www.merriam-webster.com/dictionary/weather



<br>

# Personal reflection on group work

- The way I contributed was
- What I learnt about group work was ... 